Project 3: Variant calling in Burbot (*Lota lota*) using `freebayes` and `mpileup`

### 2021-03-18 manuscript.Rmd
header change

---
title: 'Project 3: Variant calling in Burbot (*Lota lota*) using `freebayes` and `mpileup`'
author: "J Moggridge"
date: "06/03/2021"
output: 
  pdf_document:
    highlight: kate
bibliography: references.bib
link-citations: true
---

@@

<!-- purpose of variant calling -->

The analysis of genetic variants (single-nucleotide polymorphisms -SNPs and insertions/deletions - indels), by high-throughput sequencing has applications in studies of phylogeny and population structure, quantitative-trait loci, genome-wide associations, among others. Such polymorphic sites are also of practical interest in medical genomics, *eg*. to discover disease-causing mutations [@cornish2015], and in agriculture, where SNPs can be used as genetic markers to guide breeding of crops [@yao2020] and livestock [@ni2015]. As the cost of sequencing has decreased greatly and computational resources have improved, these techniques have become increasingly popular.

@@

Data can be found on `cedar` in `/scratch/jmoggrid/Project3/`. <!-- (chmod 755 /scratch/username/directory/ -R). --> <!-- srun --pty --account="def-nricker" -t 0-03:00:00 --mem=32000 /bin/bash -->

@@

I used data from project \#2 in this work, specifically the reads mapped to the burbot reference with the `bowtie2` aligner. I used `bcftools mpileup` and `GATK HaplotypeCaller` to call variants for the ten individuals using these alignments.

I first needed to convert my `.sam` files to `.bam` format and index them, as well as adding read groups using `picard` (necessary for `freebayes`).

>>>>

---
title: ''
author: ""
output: 
  pdf_document:
    highlight: kate
bibliography: references.bib
link-citations: true
---

# Project 3: Variant calling in Burbot (*Lota lota*) using `freebayes` and `mpileup`

Jason Moggridge 1159229

@@

The analysis of genetic variants (single-nucleotide polymorphisms -SNPs and insertions/deletions - indels) by high-throughput sequencing has applications in studies of phylogeny and population structure, quantitative-trait loci, genome-wide associations, among others. Such polymorphic sites are also of practical interest in clinical genomics, *eg*. to discover disease-causing mutations [@cornish2015], and in agriculture, where SNPs can be used as genetic markers to guide breeding of crops [@yao2020] and livestock [@ni2015]. As the cost of sequencing has decreased greatly and computational resources have improved, these techniques have become increasingly popular.

@@

@@

*All files can be found on `cedar` in `/scratch/jmoggrid/Project3/`.*

I used data from project \#2 in this work, specifically the reads mapped to the burbot reference with the `bowtie2` aligner. I used `bcftools mpileup` and `GATK HaplotypeCaller` to call variants for the ten individuals using these alignments. I first needed to convert my `.sam` files to `.bam` format and index them, as well as adding read groups using `picard` (necessary for `freebayes`).

<!-- fyi - bash code chunks have engine='bash' in their header -->
